---
title: "Lab 1: Intro to R"
author: ""
date: "Date of lab session"
output: oilabs::lab_report
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openintro)
library(oilabs)
```

* * *

```{r}
data(arbuthnot)
glimpse(arbuthnot)
```


### Exercise 1: 

```{r eval = FALSE}
arbuthnot$girls

```


### Exercise 2:
As per the Plot: 
Overall there is a positive trend between girls baptized over the years from 1629 to 1710.
Where their is a negative trend initially from 1629 to 1631,a rapid negative trend from 1640 to 1648 and 1702 to 1704.
A rapid positive trend can be observed from 1658 to 1663 and 1704 to 1706.


```{r}
ggplot(data = arbuthnot, aes(x = year, y = girls)) +
  geom_line()
```


### Exercise 3:
the plot shows that the proportion of boys born over time is continuously changing and it peaked in 1661

```{r echo = FALSE}
arbuthnot <- arbuthnot %>%
  mutate(Overall = boys + girls)
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / Overall)
arbuthnot
ggplot(data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_line()
```


### Exercise 4:

years included are 1940 to 2002, dimensions of the data frame is 63 rows and 3 columns and the variable names are year, boys, girls.

```{r}
present
present$year
glimpse(present)
```


### Exercise 5:
In general, the number of girls and boys from present data frame are higher than arbuthnot data frame. The magnitude of present data frame and arbuthnot data frame is not same.
```{r}
arbuthnot$girls / present$girls
arbuthnot$boys / present$boys

```

### Exercise 6:
From the plot it is evident that the proportion of boys born from 1940 to 2002 is a negative trend (decreasing) and peaked in 1946.
Yes, the arbuthnot's observation about boys being born in greater proportion than girls in the US holds true.

```{r}
present <- present %>%
  mutate(Overall = boys + girls)
present <- present %>%
  mutate(boy_ratio = boys / Overall)
present
ggplot(data = present, aes(x = year, y = boy_ratio)) + 
  geom_line()
```


### Exercise 7:
1961, the year with most total number of births in the U.S.


```{r}
present
present %>%
  arrange(desc(Overall))


```

