---
title: "Lab 3: Probability"
author: ""
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r}
library(tidyverse)
library(openintro)
```
```{r}
set.seed(42)
glimpse(kobe_basket)
kobe_basket <- kobe_basket
```

### Exercise 1
The streak length means that the number of Hit(H) occur before getting a Miss (M) in the sequence of occurences that are separated by | to signify it as a streak. The streak length of 1 means that there is 1 Hit(H) in a streak that is before getting a Miss(M) that is there are 1 (H) and 1 (M) in the streak length of 1. In streak length of 0, there are no Hit(H) before getting a Miss (M) that is there are 0 (H) and 1 (M). 

From the bar graph, streak length of 0 is the highest and streak length of 4 is the lowest.


```{r}
set.seed(42)
kobe_streak <- calc_streak(kobe_basket$shot)
ggplot(kobe_streak, aes(x=length)) +
  geom_bar()
```


### Exercise 2

From the above bar graph(Exercise 1), it can be highlighted that the most number of data points fall in the range of 0 to 1. It can be said to be a positive/right skewed distribution. The most occurred streak length was 0 and the longest streak length occurred is 4.

median = 0

```{r}
set.seed(42)
kobe_streak %>%
  summarise(Median = median(length),
            Min = min(length),
            Max = max(length))
table(kobe_streak)
coin_outcomes <- c("heads", "tails")
sample(coin_outcomes, size = 1, replace = TRUE)
sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)
sim_fair_coin
table(sim_fair_coin)
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE,
                        prob = c(0.2, 0.8))

```


### Exercise 3

In my simulation of flipping the unfair coin 100 times, i have got:
heads for 23 times and tails for 77 times.

```{r}
set.seed(42)
sim_unfair_coin1 <- sample(coin_outcomes, size = 100, replace = TRUE,
                        prob = c(0.2, 0.8))
sim_unfair_coin1
table(sim_unfair_coin1)
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 1, replace = TRUE)

```


### Exercise 4

In the sample function we have to add a probability argument with the probability of hitting should be kept as 45% and change the size to 133. as mentioned below sample(shot_outcomes, size = 133, replace = TRUE, prob = c(0.45, 0.55)

The number of times hit and miss occurred:  H 69, M 64

```{r}
set.seed(42)
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE,
                     prob = c(0.45, 0.55))
table(sim_basket)
```


### Exercise 5

sim_streak = 65*1 (65 rows and 1 column)

```{r}
set.seed(42)
sim_streak <- calc_streak(sim_basket)
sim_streak
```


### Exercise 6

From the bar graph, the distribution is looking in similar with kobe_streak data frame. 
The graph is positive/right skewed distribution. the streak length 0 occurred for the most number of times and the longest streak length is 5.

```{r}
set.seed(42)
ggplot(sim_streak, aes(x=length)) +
  geom_bar()


```


### Exercise 7

when i run the simulation of the independent shooter for a second time with different set.seed() value, I could see that both the distributions are somewhat similar. They are unimodal right skew/positively skewed. 
both the distributions have same mode as 0 (most frequent), same longest streak length as 5.

```{r}
set.seed(500)
ggplot(sim_streak, aes(x=length)) +
  geom_bar()

```


### Exercise 8

For kobe bryant distribution: 
Median: 0
Mean: 0
Max: 4
Mean: 0.7631

For simulated shooter distribution:
median: 0
min: 0
max: 5
mean: 1.0615

i think hot hand model fits with kobe's shooting patterns.
```{r}
set.seed(42)
kobe_streak %>%
  summarise(Median_kobe = median(length),
            Min_kobe = min(length),
            Max_kobe = max(length),
            Mean_kobe = mean(length))
sim_streak %>%
  summarise(Median_sim = median(length),
            Min_sim = min(length),
            Max_sim = max(length),
            Mean_sim = mean(length))
```

