---
title: "Module 11.1 Lab: Decomposition"
author: ""
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r}
library(tidyverse)
library(openintro)
library(fpp3)
library(latex2exp)
```


### Exercise 1


```{r}

global_economy %>%
  autoplot(GDP / Population, show.legend =  FALSE) +
  labs(title= "Gross Domestic Product (GDP) per capita", y = "United States($)")

```


From the plot, it is clearly demonstrating that the GDP per capita from the data set of gobal_economy is without a legend. Here, the plot is representing an upward trend. However, the effects of inflation are not taken into consideration. 


```{r}

global_economy %>%
  filter(Year == "2017") %>%
  mutate(Gdp_per_pop = GDP/Population) %>%
  arrange(desc(Gdp_per_pop))

```

In the year of 2017, which is the last year that the data from global_economy dataset possess. From the results, Luxembourg have the highest GDP per capita.

```{r}

global_economy %>%
  filter(Year == "2014") %>%
  mutate(Gdp_per_pop = GDP/Population) %>%
  arrange(desc(Gdp_per_pop))

```

1. With regards to the countries Monaco and Liechtenstein, as the data for these countries are not available for the year 2017, even after they are possessing the highest GDP per capita from the year of 1985 approximately.

2. While closely observing the results, it can be understood that the data points from the plot indicates that the country Monaco is having the highest GDP per capita of 185,152.5272 in year 2014.

3. From the results, it is showing that the country Monaco is holding the highest GDP per capita from the year 1970, barring only in three occassions/years.


### Exercise 2


A)United States GDP from global_economy.

```{r}

global_economy %>%
  filter(Country == "United States") %>%
  autoplot(GDP/Population) +
  labs(title= "Gross Domestic Product (GDP) per capita", y = "United States($)")

```


Applying a population adjustment helps to mitigate the impact of the total population on the United States' GDP, thereby offering an even more accurate depiction of the GDP's annual worth.

B) Slaughter of Victorian “Bulls, bullocks and steers” in aus_livestock.

```{r}

aus_livestock %>%
  filter(Animal == "Bulls, bullocks and steers" &
           State == "Victoria") %>%
  mutate(daily_avg_by_month = Count / days_in_month(Month)) %>%
  autoplot(daily_avg_by_month) +
  labs(title= "Slaughter of Victorian Bulls, bullocks and steers", y = "Monthly average of daily")

```


Instead of utilizing the unadjusted monthly totals, the "days_in_month" function from the lubridate library was used to modify the calendar and calculate the daily average for each month. This modification successfully accounts for the differences in the number of days in every month.


C) Victorian Electricity Demand from vic_elec.


```{r}
victorian_tib <- as_tibble(vic_elec) %>% 
  select(-c(Time))
victorian_tsib <- victorian_tib %>%
  group_by(Date) %>%
  summarise(DailyTotal = sum(Demand)) %>%
  mutate(Date = as_date(Date)) %>%
  as_tsibble(index = Date, key = DailyTotal)
victorian_tsib %>%
  ggplot(aes(x = Date, y = DailyTotal)) +
  geom_line() +
  labs(title = "Demand for Electricity",
       subtitle = "Victoria, Australia",
       y = "Total quantity (in megawatts) during the day")
```


Here, I have aimed to apply a calendar adjustment through aggregation of the daily demand totals for each date, as opposed to using the default 30-minute intervals.  From the plot it is evident that the data seemed complicated since there were so many intervals on one chart, and the small increments meant that the daily cycle, which lasted for three years, did not provide me with any useful information. I decided to combine all of the increments for a particular day in order to provide a more contextualized view, and I generated a plot that shows the overall daily demand over a three-year period. This revised plot, in my opinion, sheds more light on the seasonal variations in electricity demand and draws attention to a peculiarity that occurred at the beginning of 2014. I believe that an abnormal weather occurrence in Victoria, Australia at that time might be the cause of this anomaly.


D)Gas production from aus_production.


```{r}

lambda <- aus_production %>%
  features(Gas, features = guerrero) %>%
  pull(lambda_guerrero)
aus_production %>%
  autoplot(box_cox(Gas, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Following the transformation, gas is produced with a lambda value of $\\lambda$ = ",
         round(lambda,2))))

```


Here, I have applied Box-Cox transformation, which is a mathematical method that aids in reducing the seasonal swings throughout the course of the whole time series and is clearly been mentioned in section 3.1 of the Hyndman textbook.


### Exercise 3

Why is a Box-Cox transformation unhelpful for the canadian_gas data?

```{r}

canadian_gas %>%
  autoplot(Volume) +
  labs(title= "Natural gas production in Canada on a monthly basis", y = "Cubic meters in billions (Bcm)")

```


```{r}

lambda <- canadian_gas %>%
  features(Volume, features = guerrero) %>%
  pull(lambda_guerrero)
canadian_gas %>%
  autoplot(box_cox(Volume, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Gas production after the application of the transformation with a lambda value of $\\lambda$ = ",
         round(lambda,2))))

```

With the exception of the years 1978–1990, when there is an increase in seasonal variation, the seasonal variation in the Canadian gas production dataset is very constant through out the series, hence the Box–Cox transformation is not especially helpful. It is clear from looking at the first untransformed plot and the second plot after the Box-Cox transformation that the transformation doesn't greatly highlight the seasonal fluctuations. Actually, it seems to lessen the effects of the significant seasonal variations that were noted between 1978 and 1990..


### Exercise 4


What Box-Cox transformation would you select for your retail data (from Exercise 7 in Section 2.10)?


```{r}

set.seed(7586290)

myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

autoplot(myseries, Turnover) +
  labs(title = "The takeaway food sector in Queensland has experienced fluctuations in revenue",
       subtitle = "Series ID (A3349767W)",
       y = "Turnover")
```


```{r}

lambda <- myseries %>%
  features(Turnover, features = guerrero) %>%
  pull(lambda_guerrero)
myseries %>%
  autoplot(box_cox(Turnover, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "There has been a change in the turnover of food services with $\\lambda$ = ",
         round(lambda,2))))

```


Utilizing the guerrero feature as demonstrated in the textbook, we determithe recommended lambda value to be 0.24, indicating the application of aned  power transformation. As evident in the plot above (after applying the transformation), the seasonal variance is now much more consistent throughout the entire series.

Here, I made sure to utilize the guerrero feature as represented in the textbook, I have considered the recommended lambda value = 0.24. After applying this transformation, the plot indicates that the seasonal variance has become more consistent throughout the series. This suggests that the transformation was successful in stabilizing the variance, which can be important in making the data suitable for further analysis, such as forecasting with time series models.


### Exercise 5


For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance. Tobacco from aus_production, Economy class passengers between Melbourne and Sydney from ansett, and Pedestrian counts at Southern Cross Station from pedestrian?

A) aus_production


```{r}

lambda <- aus_production %>%
  features(Tobacco, features = guerrero) %>%
  pull(lambda_guerrero)
aus_production %>%
  autoplot(box_cox(Tobacco, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Modified the production of tobacco through $\\lambda$ = ",
         round(lambda,2))))

```


B) ansett


```{r}

lambda <- ansett %>%
  filter(Airports == 'MEL-SYD' &
           Class == 'Economy') %>%
  features(Passengers, features = guerrero) %>%
  pull(lambda_guerrero)

ansett %>%
  filter(Airports == 'MEL-SYD' &
           Class == 'Economy') %>%
  autoplot(box_cox(Passengers, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Enhanced the economy passenger experience btw Melbourne and Sydney with $\\lambda$ = ",
         round(lambda,2))))

```


C) pedestrian


```{r}

lambda <- pedestrian %>%
  filter(Sensor == 'Southern Cross Station') %>%
  features(Count, features = guerrero) %>%
  pull(lambda_guerrero)

pedestrian %>%
  filter(Sensor == 'Southern Cross Station') %>%
  autoplot(box_cox(Count, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Revolutionized pedestrian count around Southern Cross Station with $\\lambda$ = ",
         round(lambda,2))))

```


I felt that the results that demonstrated above was a bit inadequate. Because the data includes hourly counts over a two-year period, I expected a weekly pattern in the seasonality. In addition, there was a noticeable decline around the end of 2015, although I was unable to identify if this was because of holidays or another unusual occurrence in Australia. Taking these observations into consideration, I calculated lambda again using the guerrero feature, but this time I limited the computation to data from 2015 before December (see code below) in order to eliminate the notable decline. To my surprise, the lambda value that was obtained from this smaller sample stayed at -0.23. I have chosen to remain with the initial lambda value of -0.23, which was determined using the complete dataset, due to the consistency of the results.


```{r}

lambda <- pedestrian %>%
  filter(Sensor == 'Southern Cross Station' &
           Date < '2015-12-01') %>%
  features(Count, features = guerrero) %>%
  pull(lambda_guerrero)

pedestrian %>%
  filter(Sensor == 'Southern Cross Station') %>%
  autoplot(box_cox(Count, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Revolutionized pedestrian count around Southern Cross Station with $\\lambda$ = ",
         round(lambda,2))))

```


### Exercise 6


Show that a  3×5 MA is equivalent to a 7-term weighted moving average with weights of 0.067, 0.133, 0.200, 0.200, 0.200, 0.133, and 0.067.

A)
The description is about a 3x5 Moving Average, which is a two-step process where you first calculate a 5-period Moving Average (MA) and then apply a 3-period MA to the results of the 5-period MA. This process results in a smoothed time series that is symmetric because it centers the average rather than lagging behind the trend as with a trailing moving average.

Steps:

If we have a time series with 7 terms.

Calculate the 5-MA as follows:
5-MA: _, _, (Y1 + Y2 + Y3 + Y4 + Y5)/5, (Y2 + Y3 + Y4 + Y5 + Y6)/5, (Y3 + Y4 + Y5 + Y6 + Y7)/5, _, _
Compute the 3-MA based on the 5-MA values:
3-MA: _, _, _, ((Y1 + Y2 + Y3 + Y4 + Y5)/5 + (Y2 + Y3 + Y4 + Y5 + Y6)/5 + (Y3 + Y4 + Y5 + Y6 + Y7)/5)/3, _, _

Now, let's calculate the specific values for the 3X5 MA:

Y1 = 1/5 * 1/3 = (1/15)
Y2 = 2 * 1/5 * 1/3 = (2/15)
Y3 = 3 * 1/5 * 1/3 = (3/15)
Y4 = 3 * 1/5 * 1/3 = (3/15)
Y5 = 3 * 1/5 * 1/3 = (3/15)
Y6 = 2 * 1/5 * 1/3 = (2/15)
Y7 = 1/5 * 1/3 = (1/15)

After computation, the values are:

0.067
0.133
0.2
0.2
0.2
0.133
0.067


### Exercise 7

Consider the last five years of the Gas data from aus_production.

```{r}

gas <- tail(aus_production, 5*4) |> select(Gas)
head(gas)

```


A) Plot the time series. Can you identify seasonal fluctuations and/or a trend-cycle?


```{r}

gas %>%
  autoplot(Gas)

```


B) Use classical_decomposition with type=multiplicative to calculate the trend-cycle and seasonal indices.


```{r}

classical_dec <- gas %>%
  model(classical_decomposition(Gas, type = "multiplicative")) %>%
  components() 

classical_dec %>%
  autoplot() +
  labs(title = "Australia's gas production using the classical multiplicative decomposition method, calculated in petajoules")

```

C) Do the results support the graphical interpretation from part a?


Apparently the trend line here demonstrates a gradual increase from the left to right, while maintaining a relatively steady level in the middle. Over the course of the five years, the seasonal indicators show an almost perfect seasonal pattern.

D)Compute and plot the seasonally adjusted data.


```{r}

classical_dec %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Gas, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "The Petajoules",
       title = "Australia's gas production") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )

```


E)Change one observation to be an outlier (e.g., add 300 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

Outlier can be seen in the beginning


```{r}

gas_OutFront <- gas
gas_OutFront$Gas[1] <- gas_OutFront$Gas[1] + 300

OutFront_classical_dec <- gas_OutFront %>%
  model(classical_decomposition(Gas, type = "multiplicative")) %>%
  components()

OutFront_classical_dec %>%
  autoplot() +
  labs(title = "Australia's gas production using the classical multiplicative decomposition method, calculated in petajoules")

```


```{r}

OutFront_classical_dec %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Gas, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "Petajoules",
       title = "Australia's gas production") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )
```


Outlier that can be viewed in the middle


```{r}
gas_OutMid <- gas
gas_OutMid$Gas[11] <- gas_OutMid$Gas[11] + 300

outlier_middle_classical_dec <- gas_OutMid %>%
  model(classical_decomposition(Gas, type = "multiplicative")) %>%
  components()

outlier_middle_classical_dec %>%
  autoplot() +
  labs(title = "Australia's gas production using the classical multiplicative decomposition method, calculated in petajoules")
```


```{r}

outlier_middle_classical_dec %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Gas, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "Petajoules",
       title = "Australia's gas production") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )
```

Outlier that is visible at the end


```{r}

gas_OutBack <- gas
gas_OutBack$Gas[20] <- gas_OutBack$Gas[20] + 300

outlier_back_classical_dec <- gas_OutBack %>%
  model(classical_decomposition(Gas, type = "multiplicative")) %>%
  components()

outlier_back_classical_dec %>%
  autoplot() +
  labs(title = "Australia's gas production using the classical multiplicative decomposition method, calculated in petajoules")

```


```{r}

outlier_back_classical_dec %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Gas, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "Petajoules",
       title = "Australia's gas production") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )
```


F)Does it make any difference if the outlier is near the end rather than in the middle of the time series?


From the plots above, it is clear that the first outlier is very noticeable and creates an abnormality in the first part of the seasonally corrected plot. Everything about the seasonally corrected plot is similar to the original plot, except for one peculiarity. The outlier at the end exhibits the similar pattern, with the abnormality now occurring near the series' end. The seasonally adjusted plot in both situations closely resembles the first seasonally adjusted plot.On the other hand, when the middle outlier is taken into account, the seasonally adjusted plot matches the original data line much more closely, almost exactly capturing the data's natural seasonal features.  This specific middle outlier seems to raise doubts about the accuracy of the seasonally adjusted computations and graphic, at least in this particular instance.


### Exercise 8

Recall your retail time series data (from Exercise 7 in Section 2.10). Decompose the series using X-11. Does it reveal any outliers, or unusual features that you had not noticed previously?

```{r}

library(seasonal)
set.seed(7586290)

myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

x_11_decomp <- myseries %>%
  model(x11 = X_13ARIMA_SEATS(Turnover ~ x11())) %>%
  components()

autoplot(x_11_decomp) +
  labs(title =
    "Using X-11 to analyze the distribution of turnover in Queensland's takeaway food services industry")
```


In fact, there is a changing pattern that is clearly visible in the seasonal variation across time. There are peaks in the seasonal chart that is indicating more turnover at one time, and these peaks are correlating with lower turnover values in the latter portions of the plot. A few outliers that are visible from the "irregular" chart are also highlighted in the plot. Overall, the trend plot is not being able to identify any significant abnormalities in the data.


### Exercise 9


Figures 3.19 and 3.20 show the result of decomposing the number of persons in the civilian labour force in Australia each month from February 1978 to August 1995.

A) Write about 3–5 sentences describing the results of the decomposition. Pay particular attention to the scales of the graphs in making your interpretation.

The decomposition results that could be found to exhibit a trend line that has the capability of matching the actual data plot. But the sizes of the "season_year" and "remainder" components show what's really intriguing about the breakdown. In my opinion as a beginner, the "season_year" chart should normally show a thinner gray bar, which would indicate a more significant role in the data. However, it is clear from this 3.19 decomposition that the "remainder" graphic has a greater impact on the data than the "season_year" component. The reason for the "remainder" plot's increased significance is the 1991–1992 economic recession, which had such a strong influence on the data that the "season_year" component was unable to fully explain the fall.

Moreover, the sub-seasonal chart (3.20) appears to reveal significant fluctuations over the time period, particularly in the months of March, July, August, November, and December.


B) Is the recession of 1991/1992 visible in the estimated components?

As highlighted in section A, there are noticeable anomalies in the data for the years 1991 and 1992, which stand out in the "remainder" aspect of the time series decomposition. These outliers have a pronounced impact on the scale of the "remainder" decomposition. While the aggregate data and the discernible trend show a modest downturn during this period, it is the "remainder" decomposition that vividly captures the steep drop occurring in those particular years.


### Exercise 10


This exercise uses the canadian_gas data (monthly Canadian gas production in billions of cubic metres, January 1960 – February 2005).

A) Plot the data using autoplot(), gg_subseries() and gg_season() to look at the effect of the changing seasonality over time.2


```{r}

canadian_gas %>%
  autoplot(Volume)+
  labs(title = "Monthly Production of Gas in Canada",
       subtitle = "autoplot()",
       y = "Cubic meters in billions")+
  theme_replace()+
  geom_line(col = "#1B89D3")

```


```{r}

canadian_gas %>%
  gg_subseries(Volume)+
  labs(title = "Monthly Production of Gas in Canada",
       subtitle = "gg_subseries()",
       y = "Cubic meters in billions")

```


```{r}

canadian_gas %>%
  gg_season(Volume)+
  labs(title = "Monthly Production of Gas in Canada",
       subtitle = "gg_season()",
       y = "Cubic meters in billions")

```



The Canadian gas production data, as shown in the time plot, clearly demonstrates a strong upward trajectory and marked seasonal patterns. This observation is supported by both the subseries plot and the seasonal plot. Typically, gas production decreases in the summer months and increases in the winter. Notably, from 1975 to 1990, there is a noticeable intensification in seasonality, attributed to the greater contrast in gas production between the summer lows and winter highs, which is depicted by the blue and green lines on the seasonal plot.


B) Do an STL decomposition of the data. You will need to choose a seasonal window to allow for the changing shape of the seasonal component.


```{r}

canadian_gas %>%
  model(
    STL(Volume ~ trend(window = 21) +
                   season(window = 13),
    robust = TRUE)) %>%
  components() %>%
  autoplot()+
  labs(title = "Canadian Gas Production Decomposition, by applying STL")

```

The results from the STL decomposition presented above reveal that the trend component accurately reflects the fundamental direction of the original data. The seasonal component shows an increase from 1975 to 1985, then a subsequent decrease, which corresponds with the trends observed in the time plot. Additionally, the remainder component remains close to zero throughout.


C) How does the seasonal shape change over time? [Hint: Try plotting the seasonal component using gg_season().]

The seasonal pattern depicted above begins relatively level and gradually strengthens over time. In 1960, the data shows no clear trend-cycle, indicating that gas production was relatively stable during this time. However, starting from 1975, a distinct trend-cycle becomes apparent, reflecting a marked rise in gas production from that point onwards.

D) Can you produce a plausible seasonally adjusted series?


```{r}

canadian_gas %>%
 model(
    STL(Volume ~ trend(window = 21) +
                   season(window = 13),
    robust = TRUE)) %>%
  components() %>%
  ggplot(aes(x = Month)) +
  geom_line(aes(y = Volume, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(title = "Canadian Gas Production Decomposition, by applying STL") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )
```


e) Compare the results with those obtained using SEATS and X-11. How are they different?


```{r}

canadian_gas %>%
  model(x11 = X_13ARIMA_SEATS(Volume ~ x11())) %>%
  components() %>%
  autoplot()+
  labs(title = "Decomposition of Canadian Gas Production with the application of X-11 Analysis")

```


```{r}

canadian_gas %>%
  model(seats = X_13ARIMA_SEATS(Volume ~ seats())) %>%
  components() %>%
autoplot() +
  labs(title ="Decomposition of Canadian Gas Production with the application of SEATS Analysis")

```



Displayed above are the results from both the SEATS and X11 decomposition methods. There is a remarkable resemblance in the trend and seasonal components obtained from these methods. Yet, there are significant differences in seasonality when contrasted with the raw data, and the variations in the seasonally adjusted series are quite subtle between the two approaches.

Additionally, it is observed that the remainder component in the SEATS decomposition is more pronounced than in the X11 method, with both of their remainders approximately equal to one. By comparison, the remainder from the STL decomposition is notably smaller, suggesting that the STL method may provide a more accurate representation of the Canadian gas production data.

